Tidy data structure to support exploration and modeling of temporal data

### 2018-12-01 main.Rmd
draft section 5.3 extensibility

## Time-based pipeline

@@

Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data. Below conveys the pieces of critical and contextual information about the data in Table \ref{tab:tb-sub}: (1) data dimension with its shorthand time interval, alongside time zone if date-times; (2) variables that constitute of the "key" with the number of series. These details aid users in understanding their data better and manipulate the data with care.

@@

It is relatively easy to determine if the data is daily or sub-daily since most statistical packages have built-in time representations for date-times and dates, for example their corresponding classes `POSIXct` and `Date` in R. However these turn out inappropriate for weekly, monthly and quarterly data. For example, the smallest time unit of monthly data is months in a year, which is equispaced points in a year. However, using the first day of the month can range from 28 to 31 days, which is irregularly spaced. A genuine year-month representation should avoid associating days with its format. Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp. This argument is also applicable to the creation of year-week and year-quarter. (coercion)

>>>>

## Time-based pipeline {#sec:timepipeline}

@@

Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data. Below conveys the pieces of critical and contextual picture about the data in Table \ref{tab:tb-sub}: (1) data dimension with its shorthand time interval, alongside time zone if date-times; (2) variables that constitute of the "key" with the number of series. These details aid users in understanding their data better and manipulate the data with care.

@@

As a fundamental infrastructure, extensibility is highly valued and one of the core design features in the **tsibble** package. Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily relies on S3 classes and methods in R [@adv-r]. It can be extended in two major aspects: custom index and new tsibble class.

Tsibble supports an extensive range of index types from numerics to nano time, but there might be custom indices used for some occasions, for example school semesters. These academic terms vary from one institute to another within an academic year, which is defined differently from a calendar year. New index to be valid (defining `index_valid()`), the bottom line is that it can be ordered from past to future. The interval regarding semesters is further outlined through `pull_interval()`. As a result, the rest software methods such as `has_gaps()`, `count_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.

The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble. For example, a fable is actually an extension to a tsibble, mentioned in Section \ref{sec:timepipeline}. A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass. First of all, this new object itself is a tsibble. It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a fable. Tsibble verbs are also S3 generics. Developers will be able to implement these verbs for the new class if needed.
